EuPPollNet: A European database of plant-pollinator networks

### 2024-04-09 Manuscript/Ms.qmd
methods ready or more or less

main_orders = c("Hymenoptera", "Diptera", "Lepidoptera", "Coleoptera")

@@

@@

The EuPPollNet database is available in both .csv and .rds formats. The .csv file contains data in long format, while the .rds file stores the data in a list structure, with networks organized within each study. The data is organized in two different files named interaction data and flower count data. The interaction file contains `r nrow(my_tibble) + 1` columns that depict where and when the plant-pollinator interaction was observed (**Table 1**). To provide a standard unit of plant-pollinator interaction across studies, each row depicts a single interaction between a plant and a pollinator species. To build a plant-pollinator network matrix at a site level, authors only need to group interactions by plant and pollinator species, site and study. In addition, information about habitat type and bioclimatic region are also provided in this file. The floral count file contains measurements of floral availability and the unit of measurement, and should be merged with the interaction file by using the column of "Flower_data_merger".

EXPLAIN WHY IS IN A SEPARATE DATASET

Only xxxx% of studies contain floral count information. The metadata file contains [...]

Prepare final metadata

@@

kable(my_tibble, booktabs = T, linesep = "\\addlinespace", caption = "\\textbf{Table 1.}") %>%
  kable_styling(full_width = F,font_size = 9) %>% row_spec(0, bold=T) %>% 
kable_styling(position = "center")

@@

@@

All plant and pollinator species names were checked and standardized according to large scale taxonomic databases. To ensure reproducibility of the workflow, we have conducted this harmonization in R with **rgbif** [@chamberlain2022] as pivotal package to check for species names and retrieve further taxonomic information (i.e., phylum, order, family and genus) from the Global Biodiversity Information Facility (GBIF). The protocol for plants and pollinators is similar but slightly different given the availability of the different taxonomic resources. For transparency, we have included in the database the old species name, the new assigned name, and, if name of the species is uncertain (e.g., species complex or species alike).

@@

bootstraping, inext package

@@

@@

Geographic and taxonomic biases!

>>>>

#Check number of species and percentage of interactions of poll. groups
main_orders = c("Hymenoptera", "Diptera", "Lepidoptera", "Coleoptera")

@@

#Check percentage of studies with flower counts
floral_counts_percent = data  %>%
select(Study_id, Flower_data) %>% 
distinct() %>% 
group_by(Flower_data) %>% 
summarise(Percentage = n()/ nrow(.)*100) %>% 
filter(Flower_data == "Yes")

@@

The EuPPollNet database is available in both .csv and .rds formats. The .csv file contains data in long format, while the .rds file stores the data in a list structure, with networks organized within each study. The file contains a total of `r nrow(my_tibble) + 1` columns that depict where and when the plant-pollinator interaction was observed (**Table 1**). To provide a standard unit of plant-pollinator interaction across studies, each row depicts a single interaction between a plant and a pollinator species. To build a plant-pollinator network matrix at a site level, authors only need to group interactions by plant and pollinator species, site and study. Information about habitat type and bioclimatic region are also provided in this file. In addition, the interaction dataset includes a column that describes the presence or absence of floral counts for each study. The flower count data is provided in a separate file given that one third of studies lack floral measurements and that the units differ greatly across studies. Finally, metadata at the study level is provided, including information about the authors, digital object identifier if available, sampling time and taxonomic coverage of the main pollinator groups for each study.

@@

kable(my_tibble, booktabs = T, linesep = "\\addlinespace", caption = "\\textbf{Table 1.} Column names and their descriptions within the EuPPollNet database.") %>%
  kable_styling(full_width = F,font_size = 9) %>% row_spec(0, bold=T) %>% 
kable_styling(position = "center")

@@

The completeness of the EuPPollNet database was evaluated by exploring the rarefied accumulation curves of plant and pollinator species and their interactions across the different sampling sites. In addition, the accumulation curve of pollinator species with increasing number plant species was also explored. The rarefied and extrapolated sampling curves were obtained with the help of the **iNEXT** package [@hsieh2016]. This was complemented with 100 random accumulation curves obtained programmatically to guide the visualization of the different rarefied curves.

@@

All plant and pollinator species names were checked and standardized according to large scale taxonomic databases. To ensure reproducibility of the workflow, we have conducted this harmonization in R with **rgbif** [@chamberlain2022] as pivotal package to check for species names and retrieve further taxonomic information (i.e., phylum, order, family and genus) from the Global Biodiversity Information Facility (GBIF). The protocol for plants and pollinators is similar but slightly different given the availability of the different taxonomic resources. For transparency, we have included in the database the old species name, the new assigned name, and, if the name of the species is uncertain (e.g., species complex or species alike).

@@

@@

## Network analyses

To get a general overview of some of the main aspects of these set of plant-pollinator networks, we evaluated how connectance and nestedness change across the latitudinal range of studies and bioclimatic regions. In addition, we explored the association between connectance and number of species per network (i.e., geometric mean of plants and pollinators), and if the different networks are more or less nested than expected by chance. Importantly, we implemented two different metrics of nestedness given the different analytical objectives. Common metrics of nestedness are highly correlated to network size, limiting our ability to compare this metric across networks. To avoid this, we implemented a normalised version of nestedness (NODFc) that corrects by connectance and the number of species [@song2017]. This metric was calculated with the help of the **maxnodf** package [@hoeppke2021], which employs a computational efficient approach to calculate it. However, to compare if networks are more or less nested than expected by chance, we employed the traditional z-score approach with the widely used nestedness metric (NODF) from @almeida2008, as this approach allow us to compare our results with previous published nestedness analyses in plant-pollinator networks and only compares each unique network against their randomized versions. We calculated 100 null models for each network with the help of the _vaznull_ function from the package **Bipartite** [@dormann2008] that implements the null model from @vazquez2005. These null networks have the same connectance and number of plant and pollinator species as the empirical ones, but different marginal totals. In other words, interactions are reshuffled with almost equal probabilities to their relative abundances. Finally, connectance was estimated for each network by using the function _networklevel_ from **Bipartite**.

@@

Geographic and taxonomic biases!
Supporting reference:
https://doi.org/10.1111/ecog.05926

Networks, ref:

https://www.biorxiv.org/content/10.1101/2020.04.02.021691v1.abstract
